---
title: "Date mischief"
output: 
  blogdown::html_page: 
    highlight: tango
    css: ../../css/camp_style.css
    number_sections: false
    self_contained: false
    fontsize: 18pt
    monofont: Source Code Pro
    toc: false
    toc_depth: 1
---


```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(lubridate)
library(palmerpenguins)
library(janitor)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Getting started

The research crew wants to determine if there is a relationship between isotope levels in the `penguins_raw` data and temperature. We have temperature data for each of the islands uploaded to the webpage.


Let's get all of this data into R so we can combine it with our `penguins_raw` data. 

```{r read_data}
library(tidyverse)

biscoe <- read_csv("https://tidy-mn.github.io/R-camp-penguins/data/Biscoe_temperatures.csv")

dream <- read_csv("https://tidy-mn.github.io/R-camp-penguins/data/Dream_temperatures.csv")

torg <- read_csv("https://tidy-mn.github.io/R-camp-penguins/data/Torgersen_temperatures.csv")
```


## Data exploration

Let's do a quick overview of our temperature data. As you are exploring the data, look for what is different between the 3 data frames? 

What can we do to make the data more tidy?

```{r explore_data, eval=F}
glimpse(biscoe)
glimpse(dream)
glimpse(torg)
```


## Clean-up time

### Step 1

Take a closer look at the `torg` data. Why does it have so many more rows? Try arranging the data by the *date* column, what do you notice?

```{r arrange}
arrange(torg, date)
```

<br>

> Icky...It looks like there are duplicate rows in the data. That seems to happen a lot. Since the temp value always appears to be the same for the duplicate row, I think it is safe to say the duplicate should be removed. 

If you are 100% sure your data should not have duplicate rows in it, you can use the function `distinct( )` to keep only one row from each of the groups of duplicates. In the end, every row will be a distinct and unique row compared to ALL the others in the table.


```{r distinct}
torg <- distinct(torg)
```


> How many rows does *torg* have now?


<br>

### Step 2 


The island names are in the file name, but are not in the data itself. That's not ideal when you're looking to join data together or perhaps want to label points on a chart with the island name.

Let's add the island names to use for joining later. Remember that R is picky about name matching, so it's important to make sure the names are exactly correct, including CAPs. 

The island names are `Biscoe`, `Dream`, and `Torgersen`. *Copy/Paste* is your friend here. Use `mutate( )` to add a new column called *island* to each of the data sets.


<div class="quiz">

### <i class="fas fa-cookie-bite" style="color: gray;"></i> Which island? {-}

```{r, results='asis', echo=F}
source("../../insert_tabs.R")

tabs <- c("Description", "Show hint", "Show code")

content <- c(Exercise = '<b> Use `mutate( )` to add a new column called *island* to each of the data sets.</b>

<br>',

show_hint = 'code_start
biscoe <- biscoe %>%
          mutate(island = ______ )

dream <- dream %>%
         
#.....
  
code_end',

show_code = 'code_start
biscoe <- biscoe %>%
          mutate(island = "Biscoe")

dream <- dream %>%
         mutate(island = "Dream")

torg <- torg %>%
        mutate(island = "Torgersen")
  
code_end')

tab_html <- add_tabs(tabs, content, "0a")

cat(paste0(tab_html, collapse = "\n"))
```

</div>

```{r add_islands, echo=F}
biscoe <- biscoe %>%
          mutate(island = "Biscoe")

dream <- dream %>%
         mutate(island = "Dream")

torg <- torg %>%
        mutate(island = "Torgersen")
```


<br>


When exploring the data, did you notice anything about the dates? They are all in different formats! Luckily we have `lubridate` to come to the rescue. Let's convert those dates so they are all in the same format. Remember to choose your date function from the table below.

<div class="bluecode">

| Format | Function to use   |
|:----------------------------------|:--------|
| Month-Day-Year ~ "<span class="teal">05-18-2023</span>" or "<span class="teal">05/18/2023</span>"                | `mdy(`date`)  `    |  
| Day-Month-Year ~ "<span class="teal">18-05-2023</span>" or "<span class="teal">18/05/2023</span>"  | `dmy()  `    | 
| Year-Month-Day ~ "<span class="teal">2023-05-18</span>" or "<span class="teal">2023/05/18</span>" | `ymd()  `    | 

</div>


<div class="quiz">

### <i class="fas fa-cookie-bite" style="color: gray;"></i> Make it a Date {-}

```{r, results='asis', echo=F}
source("../../insert_tabs.R")

tabs <- c("Description", "Show hint", "Show code")

content <- c(Exercise = '<b> Use `mutate( )` to convert the *date* column to a Date object in each of the data sets.</b>

<br>',

show_hint = 'code_start

biscoe <- biscoe %>%
          mutate(date = mdy(date))

dream <- dream %>%
         mutate(date = ________ )
         
#.....
  
code_end',

show_code = 'code_start
biscoe <- biscoe %>%
          mutate(date = mdy(date))

dream <- dream %>%
         mutate(date = dmy(date))

torg <- torg %>%
        mutate(date = ymd(date))
  
code_end')

tab_html <- add_tabs(tabs, content)

cat(paste0(tab_html, collapse = "\n"))
```

</div>


```{r convert_dates, echo=F}
biscoe <- biscoe %>%
          mutate(date = mdy(date))

dream <- dream %>%
         mutate(date = dmy(date))

torg <- torg %>%
        mutate(date = ymd(date))
```


## Binding them all together

To join the temperature data to our penguin data easily, we want to combine all 3 data frames into one data frame. Keeping in mind tidy data, what is the best way to do this?

```{r combine}
temps <- bind_rows(biscoe, dream, torg)
```


Run a few tests on the new table to ensure all the island data survived the `bind_rows` journey.

```{r, eval=F}
# Check that all the islands made it into temps table
distinct(temps, island)

# Does the total number of rows check out
nrow(temps)

nrow(biscoe) + nrow(dream) + nrow(torg)
```



## More data exploartion

We have our temperature data in a happy place, but what about the penguin data? 
Let's get the data and determine which columns we can use to combine with our temperature data.

```{r explore_penguin_data}
library(palmerpenguins)

penguins_raw <- penguins_raw

glimpse(penguins_raw)
```


<br>

Both data frames have an **island** column and a **date** column, but with different names. What operation do we want to use? Which join type is most appropriate here? Remember that when combining data frames, column names are important and are case-sensitive. We want the resulting data frame to have the same number of rows as the *penguins_raw* data frame.


<div class="quiz">

### <i class="fas fa-cookie-bite" style="color: gray;"></i> Table join {-}

```{r, results='asis', echo=F}
source("../../insert_tabs.R")

tabs <- c("Description", "Show hint", "Show code")

content <- c(Exercise = '<b> Use a `****_join( )` function to combine the columns of the two data sets so the result has the same number of rows as *penguins_raw*.</b>

<br>',

show_hint = 'code_start
penguin_temps <- left_join(penguins_raw, temps,
                           by = c("Island" = "_______" ,
                                  "Date Egg" =  "______" ))
  
# inner_join() would also work in this case

code_end',

show_code = 'code_start
penguin_temps <- left_join(penguins_raw, temps,
                           by = c("Island" = "island",
                                  "Date Egg" = "date"))
  
code_end')

tab_html <- add_tabs(tabs, content)

cat(paste0(tab_html, collapse = "\n"))
```

</div>


```{r join, echo=F}
penguin_temps <- left_join(penguins_raw, temps,
                            by = c("Island" = "island",
                                   "Date Egg" = "date"))
```



## The relationship between temp and isotopes

With all of the data in one data frame, we can now plot isotope levels versus temperatures. But first, we need to clean up those long tricky column names. How do we do that?

```{r tricky-names}
# The current names
names(penguin_temps)
```


```{r clean_names}
#install.packages("janitor")
library(janitor)

penguin_temps <- clean_names(penguin_temps)

names(penguin_temps)
```


## Plots

We can finally make our plots. We're interested in temperatures in comparison to isotope levels. These were in the columns `Delta 15 N (o/oo)` and `Delta 13 C (o/oo)` in the original *penguins_raw* data frame before the column names were cleaned. Which type of plot do you think is most appropriate here? Add some color to show the differences between the 3 islands.


```{r plots, echo=F, out.width="73%"}
# Delta 15 plot
ggplot(penguin_temps,
       aes(x = temperature_c, y = delta_15_n_o_oo, color = island)) +
  geom_point(size = 3)

# Delta 13 plot
#ggplot(penguin_temps,
#       aes(x = temperature_c, y = delta_13_c_o_oo, color = island)) +
#  geom_point(size = 3) +
#  scale_color_discrete(type = c("green", "blue", "purple"))
```


<div class="quiz">

### <i class="fas fa-cookie-bite" style="color: gray;"></i> Isotope plots {-}

```{r, results='asis', echo=F}
source("../../insert_tabs.R")

tabs <- c("Description", "Show hint", "Show code")

content <- c(Exercise = '<b> Make 2 plots. One comparing temperature to Delta 15 N, and one comparing temperature to Delta 13 C. Assign the data from each island to its own color.</b>

<br>',

show_hint = 'code_start
# Delta 15 plot
ggplot(penguin_temps,
       aes(x = temperature_c, y = delta_15_n_o_oo, _______ = _______ )) +
  geom_*****(size = 3)

# Delta 13 plot
ggplot(penguin_temps,
       aes(x = temperature_c, y = ......

code_end',

show_code = 'code_start
# Delta 15 plot
ggplot(penguin_temps,
       aes(x = temperature_c, y = delta_15_n_o_oo, color = island)) +
  geom_point(size = 3)


# Delta 13 plot
ggplot(penguin_temps,
       aes(x = temperature_c, y = delta_13_c_o_oo, color = island)) +
  geom_point(size = 3) +
  scale_color_discrete(type = c("green", "blue", "purple"))
  
code_end')

tab_html <- add_tabs(tabs, content)

cat(paste0(tab_html, collapse = "\n"))
```

</div>



## Expedition complete

**Congratulations!** You have learned a bit about how temperature may affect penguin chemistry. 

Time to prep for the next expedition...

<br>

<i class="fa fa-cookie-bite"></i> 
<i class="fa fa-robot"></i> 
<i class="fa fa-cookie-bite"></i> 
