---
title: "My Penguin Report"
author: "Barbara Monaco"
date: "`r lubridate::today()`"
format: 
  html: default
  pdf: default
editor: visual
---

# Introduction

This is good place to talk about what you want people to know about the work that you're doing. What data are you using? What does it mean? What questions are you trying to answer and why?

## Setup 

What do I need to make it all work?

```{r}
#| label: load-packages
#| warning: false

library(tidyverse)
library(palmerpenguins)
library(janitor)
library(gt)
```

## Read in data

Get the data into R because that's important.

```{r}
data(penguins)
```


## Cleaning data

All the steps to make the data more usable, because strangely...it usually isn't.

```{r}
penguins_raw_clean <- penguins_raw %>% 
  clean_names()
```

# Graphs and summaries

Ah yeah, let's see that good stuff - pretty pictures and amazing tables will make your report stand out from the crowd!

```{r}
tabyl(penguins_raw_clean, species, island) %>% 
  adorn_totals(where = 'row') %>% 
  adorn_totals(where = 'col')
```

```{r}

penguin_island_dist <- penguins_raw_clean %>% 
  group_by(region, island, species) %>% 
  summarise(count = n())

ggplot(data = penguin_island_dist, 
       aes(x = island, y = count, fill = species)) +
  geom_col(position = position_dodge2(preserve = "single")) + 
  geom_text(aes(label = count, y = count + 5), 
            position = position_dodge2(width = 0.9)) +
  scale_fill_brewer(type = "qual", palette = 3) +
  labs(title = "Distribution of penguin species by island",
       x = "Island",
       y = "Number of penguins observed") +
  theme_bw()

```

# Conclusions and Next Steps

Have any thoughts about what you'd do next if you were the analytic bot on the research vessel looking for cool penguin data? Or what you might do next if you were you working on your own project?

## Lessons learned

Some things I learned:

> I learned how to load in libraries, use functions, and make data bend to my every whim (insert evil laugh here). I also learned how to make awesome graphs, maps, and tables - oh my!

## What I liked best about R Camp

Here's what I liked best and your thing might be different from my thing and that's cool.

> I liked making videos and hanging out with my friends - we don't always get to work together so it's extra nice when we do!

> I also enjoyed making new friends...Hello NEW FRIEND!
