---
title: "ggplot exercises"
output: 
  blogdown::html_page: 
    highlight: tango
    css: "../../css/camp_style.css"
    number_sections: false
    self_contained: false
    fontsize: 18pt
    monofont: Source Code Pro
    toc: true
    toc_depth: 1
---

# Palmer Penguin Exploration - Part 1

We met some very excited researchers who wanted to share their data with us when they heard we really liked penguins. They also were hoping that maybe we would look at their data and help them answer some questions about it. What do you say? Are you willing to help? 

THANK YOU! Here's the questions the researchers were hoping to have answered - 

- How can we determine the distribution of species by island?
- Which species has the longest bill?
- What measurements are related to each other? For example is a penguin with a long bill more likely to have a larger body mass? Also what relationship does species have to measurements?

Let's make a graphical analysis of the data for them!

```{r, message = FALSE, warning= FALSE, class.source = 'fold-show'}
library(tidyverse)
# install.packages('palmerpenguins') #you might not have this package yet so make sure you install it!
library(palmerpenguins)

data("penguins") #this will read in the data from the package and load it to to our environment
```

## What does the data look like?

As good scientists we should always start with looking at the data provided to us and understanding what information was collected. What are some options for just taking a quick *peek* at the data?

```{r}
glimpse(penguins)
summary(penguins)
```

What do we know that we didn't before after running these commands? We know that there are 8 columns, three of them are factors and the rest are numeric (either integer or double).

We also know that Chinstrap are the least prevalent and Biscoe Island has the most penguins. On a more unhappy note, we also see that there are missing data, some for the measurement data but most are for sex (11 NA's).

Let's go back to the researchers' questions.

## Species by Island


```{r}
ggplot(data = penguins, aes(x = island, fill = species)) +
  geom_bar() 
```

Great! We have a stacked bar chart that shows the species by island! But the researchers have asked us to make it look "nicer" - things like better labels, colors, etc.

Back to the drawing board...let's start with labels! We want to add a title and make the axis labels be capitalized vs lower-case. Also someone complained about the background color and asked us to make it look "cleaner".

```{r}
ggplot(data = penguins, aes(x = island, fill = species)) +
  geom_bar() +
  xlab("Island") + 
  ylab("Number of Penguins observed") +
  labs(title = "Distribution of penguin species by island") +
  theme_bw()
```

Okay - whew! We added a title, a nice x-axis and y-axis label, changed the theme to `theme_bw` which we think looks pretty clean so hopefully they do too.

Now we just need to tackle color...hmmm, we know it's a bar-chart which means that we need to specify 'fill' vs 'color' so we can use the function `scale_fill_brewer` to pick out our favorite (RColorBrewer palette)[https://colorbrewer2.org/] for this graph. We want a qualitative one and we also want one that is colorblind safe so let's pick palette 3!

```{r}
ggplot(data = penguins, aes(x = island, fill = species)) +
  geom_bar() +
  scale_fill_brewer(type = 'qual', palette = 3) +
  xlab("Island") + 
  ylab("Number of Penguins observed") +
  labs(title = "Distribution of penguin species by island") +
  theme_bw()
```

Annnnd of course one the researchers had a last minute request. They didn't like the stacked bar chart and they wanted to know if we could make them side-by-side instead. As an extra challenge see if you can figure out how to make each bar the same width as well!

```{r}
ggplot(data = penguins, aes(x = island, fill = species)) +
  geom_bar(position = position_dodge2(preserve = 'single')) +
  scale_fill_brewer(type = 'qual', palette = 3) +
  xlab("Island") + 
  ylab("Number of Penguins observed") +
  labs(title = "Distribution of penguin species by island") +
  theme_bw()
```

## Which species has the longest bill?

```{r}
ggplot(data = penguins, aes(x = bill_length_mm))+
  geom_histogram(aes(fill = species),
                     alpha = 0.5, 
                     position = 'identity') + # need to have position = 'identity' to have the alpha work!
  scale_fill_brewer(type = 'qual', palette = 3) +
  xlab("Bill Length (mm)") + 
  ylab("Number of Penguins observed") +
  labs(title = "Distribution of bill length (mm) by species") +
  theme_bw()
```

Looks like Gentoo and Chinstrap are pretty much tied for longest bill! Wonder if bill length is related to other measurements...Always more questions!

## What measurements are related to each other?

```{r}
cor(penguins[3:6])
```

Why are we getting NAs? Is it because of the missing values? Okay fine, let's only use complete observations.

```{r}
cor(penguins[3:6], use = 'complete.obs')
```

Excellent! So body mass and flipper length are VERY correlated and flipper length and bill length and depth also seem to be related. Honestly a lot of these seem to related so let's go for the MOST correlated and take a look - and remember to bring in species!

```{r}
ggplot(data = penguins, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color = species,
             shape = species))+
  scale_color_brewer(type = 'qual', palette = 3) +
  xlab("Body Mass (g)") + 
  ylab("Flipper Length (mm)") +
  labs(title = "Relationship of body mass and flipper length by species") +
  theme_bw()
```

Fascinating...Gentro is really just all in a class by themselves. Otherwise it looks like the overall relationship between body mass and flipper length is the same amoung the different penguin species. You can add a trend-line to the graph to see that easier or see if the human eye is being tricked by very fun looking dots.

```{r}
ggplot(data = penguins, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color = species,
             shape = species))+
  geom_smooth(method = 'lm', se = FALSE, aes(color = species)) + 
  scale_color_brewer(type = 'qual', palette = 3) +
  xlab("Body Mass (g)") + 
  ylab("Flipper Length (mm)") +
  labs(title = "Relationship of body mass and flipper length by species") +
  theme_bw()
```

Let's try doing the same exercise but let's use the `facet_wrap` function.

```{r}
ggplot(data = penguins, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color = species,
             shape = species))+
  geom_smooth(method = 'lm', se = FALSE, aes(color = species)) + 
  scale_color_brewer(type = 'qual', palette = 3) +
  xlab("Body Mass (g)") + 
  ylab("Flipper Length (mm)") +
  labs(title = "Relationship of body mass and flipper length by species") +
  theme_bw() +
  facet_wrap(facets = ~species)
```

What's the danger in using the `scales = 'free'` argument? What happens when you add it? How would you explain the result to someone else?

As we've said before...always more questions!

Try to think of your own question that you use this data to answer or go back to the correlation matrix and see if there's other relationships you'd like to explore!

The researchers have one more thing they'd like to say...

# THANK YOU! {-}
