---
title: "Inventory Duty - Exercise 1"
output: 
  blogdown::html_page: 
    highlight: tango
    css: ../../css/camp_style.css
    number_sections: true
    self_contained: false
    fontsize: 18pt
    monofont: Source Code Pro
    toc: false
    toc_depth: 1
---

<link rel="stylesheet" 
href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/6.4.2/css/all.min.css">

<style>
h1 .header-section-number {background: cornflowerblue;}
hr {background:  cornflowerblue;}
</style>


```{r setup, include=FALSE}
library(tidyverse)
library(rmarkdown)
library(knitr)
library(magrittr)
library(icons)

xaringanExtra::use_clipboard()

opts_chunk$set(error=TRUE, color=NULL, message=FALSE, eval=T)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
```

# Before you start...

Please make a brand-new project for this exercise. This will be our exercise project where we'll do all our exercises for the remainder of the course.  


- To make a new project: Go to `File` -> `New Project` -> `New Directory` -> `New Project`

Pick a new directory for your project to live. Choose a location that makes sense to you so you'll be able to find it again. If you don't change the location, the default setting will be used and your project will be created in your `Documents/` folder. 

Click **Create project**.

Now that you have a new project setup (you should see your project name in the top right of RStudio), create a new R script and save it. 

- To make a new R script: 
    - Go to `File` -> `New File` -> `R Script` -or- 
    - Click the top left icon with the green plus sign
 

You can now save the script by clicking <kbd>CTRL</kbd> + <kbd>S</kbd> or by clicking the floppy disk icon. Name your script something pithy like `ex_1_day_1.R` or `frankie.R`. 

You're ready to go!


# Inventory duty

We love our job as baking-bot, but the thing we love more than anything is taking inventory of ingredients. We like to play with the inventory data and present beautiful summaries to the lead baking-bot.  

The bakery is expecting some big orders for the upcoming holidays and wants to be prepared. We've been asked to download the current inventory and determine if we are running low on any ingredients.

Let's read in the inventory data!

```{r}
inventory <- read_csv("https://tidy-mn.github.io/R-camp-penguins/data/inventory_data.csv")
```


# Getting to know the data

Now that we have the data let's take some steps to better understand the information in the file. 

- How many rows are there, how many columns? 
- What data types are the columns? 
- What's the general range of the values? 
- â­ The lead baking-bot especially wants to know what the maximum value is for `NUMBER_OF_UNITS_IN_STOCK` for very important planning purposes. 

Feel free to use any and all of the functions you know, but also feel free to use the hints below for inspiration!

<div class="quiz">

### <i class="fas fa-cookie-bite" style="color: gray;"></i> Get to know the data {-}

```{r exer-tabs1, results='asis', echo=F}
source("../../insert_tabs.R")

tabs <- c("Explore functions", "Show hint", "Show code")

content <- c(Exercise = '<b> When we first get a dataset it is fun to sit down and get to know it better. Below are a few functions to choose from: </b>
  
- `glimpse`
- `summmary`
- `head`
- `tail`
- `class`

<br>',

show_hint = 'code_start
# glimpse() gives us that data type for each column and 
# a few sample values for each column

glimpse() 

# summary() shows us missing values and stats for numeric columns

summary() 

code_end',

show_code = 'code_start
glimpse(inventory)

summary(inventory)

code_end')

tab_html <- add_tabs(tabs, content, 1)

cat(paste0(tab_html, collapse = "\n"))
```
</div>

<br>

WHOA. We have 378 units of something in stock?! What are we making, cupcakes for a whole school?! A cake big enough to cover a bus?! Time to go investigate...


# Filtering to the good stuff

How do we go about trying to figure out what's going on with that wacky number? We need to be able to narrow down just to the row that meets the condition where `NUMBER_OF_UNITS_IN_STOCK` is equal to 378. What tool do we have that will let us do that?

<div class="quiz">

### <i class="fas fa-cookie-bite" style="color: gray;"></i> Investigation underway {-}

```{r exer-tabs2, results='asis', echo=F}

tabs <- c("Find the odd-ball", "Show hint", "Show code")

content <- c(Exercise = '<b> Filter the inventory data to just the row that meets the condition of `NUMBER_OF_UNITS_IN_STOCK` is equal to 378 </b>

<br>',

show_hint = 'code_start
# Remember to use the double equal sign (==) for "is equal to"

filter(inventory, PLACE A CONDITION HERE) 

code_end',

show_code = 'code_start

filter(inventory, NUMBER_OF_UNITS_IN_STOCK == 378)

code_end')

tab_html <- add_tabs(tabs, content, 2)

cat(paste0(tab_html, collapse = "\n"))
```
</div>

<br>

Look at that. We have a **total** row - that would a wrench in our summaries. That row is also adding `NA`s to other columns which might make some of our analysis harder in the future. Let's get rid of this summary row and keep only the *real data*.

> **NOTE:**  Lots of functions you'll meet in R will have the additional argument `na.rm = TRUE`, which allows us to ignore NA values when calculating things like means, maxes, etc. Super handy!


<div class="quiz">

### <i class="fas fa-cookie-bite" style="color: gray;"></i> Filter time {-}

```{r exer-tabs22, results='asis', echo=F}

tabs <- c("Filter out", "Show hint", "Show code")

content <- c(Exercise = '<b> Filter the inventory data to drop the TOTAL row. </b>

<br>',

show_hint = 'code_start
# Remember the not equal sign is "!="

inventory <- filter(inventory, PLACE A CONDITION HERE)

code_end',

show_code = 'code_start

inventory <- filter(inventory, INGREDIENT != "TOTAL")

code_end')

tab_html <- add_tabs(tabs, content, 55)

cat(paste0(tab_html, collapse = "\n"))
```
</div>


Let's try `summary` again and see if we can get some better results, hopefully no more NAs or outrageous numbers.

```{r}
inventory <- filter(inventory, INGREDIENT != "TOTAL")

summary(inventory)
```

Much better. Okay, let's see what we've been asked to do.


# Place that order!

We need to use the data we received to determine what items to order. If the number of units we currently have in-stock is less than the order threshold, we need to place an order right away.

- Do we have any ingredients (rows in the data) that meet this criteria?

<div class="quiz">

### <i class="fas fa-cookie-bite" style="color: gray;"></i> Time to order? {-}

```{r exer-tabs3, results='asis', echo=F}
source("../../insert_tabs.R")

tabs <- c("Less than Order Threshold", "Show hint", "Show code")

content <- c(Exercise = '<b> Filter the inventory data to the rows that meet the condition where the number of units in stock is less than the order threshold. </b>

<br>',

show_hint = 'code_start
# You will need to compare 2 columns in the data

order_now <- filter(inventory, PLACE A CONDITION HERE)

order_now

code_end',

show_code = 'code_start
order_now <- filter(inventory, 
                    NUMBER_OF_UNITS_IN_STOCK < ORDER_THRESHOLD)
                    
order_now

code_end')

tab_html <- add_tabs(tabs, content, 32)

cat(paste0(tab_html, collapse = "\n"))
```
</div>

<br>

That's just what I expected. We've been getting lots of orders for our **Gluten-free Applesauce** and **Walnut muffins**, and also for our **Triple-Chocolate Brownies**. Time for an urgent supply order so we don't run out!

The last thing to do is to write/save the filtered data. 

```{r, eval=F}
write_csv(x = order_now, file = "order_now.csv")
```


<br>

Excellent! We'll send this out so we get those ingredients right away. Even better, next inventory time we'll have these scripted steps that we can use again and again and again...well you get the idea.

__NICE WORK!__

<i class="fa fa-cookie-bite"></i> 
<i class="fa fa-robot"></i> 
<i class="fa fa-cookie-bite"></i> 
